### 2024-04-16 paper/paper.qmd
Created graphs for data section in quarto document

---
title: "Is it Better to be a Moose or a Black Bear?"
subtitle: "An Analysis of Ontario Wildlife Harvest and Active Hunter Rates From 2012 to 2023"
author: Samantha Barfoot
thanks: "Code and data are available at: https://github.com/SamanthaBarfoot/wildlife_hunting.git"
date: today
date-format: long
abstract: "This paper analyzes Ontario wildlife harvest counts, specifically moose and black bears, and the number fo active hunters based on region. The analysis examines harvest counts from 2012 to 2023 across northern, southeastern, and southwestern Ontario. It was found that..."
format: pdf
number-sections: true
table-of-contents: true
bibliography: references.bib
---

```{r}
#| echo: false
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
```

# Introduction

The analysis, graphs, and tables were made with [@citeR].

# Data
```{r}
#| echo: false
#| include: false
#| warning: false
#| message: false
#### Read in data ####
north_bear <-
  read_csv(
    file = "../data/analysis_data/north_data_bear.csv",
    show_col_types = FALSE
  )
se_bear <-
  read_csv(
    file = "../data/analysis_data/southeast_data_bear.csv",
    show_col_types = FALSE
  )

sw_bear <-
  read_csv(
    file = "../data/analysis_data/southwest_data_bear.csv",
    show_col_types = FALSE
  )

north_moose <-
  read_csv(
    file = "../data/analysis_data/north_data_moose.csv",
    show_col_types = FALSE
  )

se_moose <-
  read_csv(
    file = "../data/analysis_data/southeast_data_moose.csv",
    show_col_types = FALSE
  )

sw_moose <-
  read_csv(
    file = "../data/analysis_data/southwest_data_moose.csv",
    show_col_types = FALSE
  )
```

```{r}
#| echo: false
#| warning: false
#| label: fig-north
#| fig-cap: "Northern harvest numbers for moose and bears between 2009 and 2023"

#create line graph 
ggplot()+
  geom_line(data = north_bear, aes(x = Year, y = Harvest), color = "blue") + 
  geom_line(data = north_moose, aes(x = Year, y = `Total Harvest`), color = "red")+
  theme_minimal()

```

```{r}
#| echo: false
#| warning: false
#| label: fig-southeast
#| fig-cap: "Southeastern harvest numbers for moose and bears between 2009 and 2023"

#create line graph 
ggplot()+
  geom_line(data = se_bear, aes(x = Year, y = Harvest), color = "blue") + 
  geom_line(data = se_moose, aes(x = Year, y = `Total Harvest`), color = "red")+
  theme_minimal()

```

```{r}
#| echo: false
#| warning: false
#| label: fig-southwest
#| fig-cap: "Southwestern harvest numbers for moose and bears between 2009 and 2023"

#create line graph 
ggplot()+
  geom_line(data = sw_bear, aes(x = Year, y = Harvest), color = "blue") + 
  geom_line(data = sw_moose, aes(x = Year, y = `Total Harvest`), color = "red")+
  theme_minimal()

```

```{r}
#| echo: false
#| warning: false
#| label: fig-bearharvest
#| fig-cap: "Bear harvest numbers for the northern, southeastern and southwestern regions from 2012-2023"

#create line graph 
ggplot()+
  geom_line(data = sw_bear, aes(x = Year, y = Harvest), color = "blue") + 
  geom_line(data = se_bear, aes(x = Year, y = Harvest), color = "red")+
  geom_line(data = north_bear, aes(x = Year, y = Harvest), color = "green")+
  theme_minimal()
```

```{r}
#| echo: false
#| warning: false
#| label: fig-mooseharvest
#| fig-cap: "Moose harvest numbers for the northern, southeastern and southwestern regions from 2012-2023"

#create line graph 
ggplot()+
  geom_line(data = sw_moose, aes(x = Year, y = `Total Harvest`), color = "blue") + 
  geom_line(data = se_moose, aes(x = Year, y = `Total Harvest`), color = "red")+
  geom_line(data = north_moose, aes(x = Year, y = `Total Harvest`), color = "green")+
  theme_minimal()
```

```{r}
#| echo: false
#| warning: false
#| label: fig-swhunters
#| fig-cap: "Number of active bear and moose hunters in the southwest region from 2009-2023"

#create line graph 
ggplot()+
  geom_line(data = sw_bear, aes(x = Year, y = `Active Hunters`), color = "blue") +
  geom_line(data = sw_moose, aes(x = Year, y = `Active Hunters`), color = "red")+
  theme_minimal()

```

```{r}
#| echo: false
#| warning: false
#| label: fig-sehunters
#| fig-cap: "Number of active bear and moose hunters in the southeast region from 2009-2012"

#create line graph 
ggplot()+
  geom_line(data = se_bear, aes(x = Year, y = `Active Hunters`), color = "blue") +
  geom_line(data = se_moose, aes(x = Year, y = `Active Hunters`), color = "red")+
  theme_minimal()

```

```{r}
#| echo: false
#| warning: false
#| label: fig-northhunters
#| fig-cap: "Number of active bear and moose hunters in the northern region from 2009-2023"

#create line graph 
ggplot()+
  geom_line(data = north_bear, aes(x = Year, y = `Active Hunters`), color = "blue") +
  geom_line(data = north_moose, aes(x = Year, y = `Active Hunters`), color = "red")+
  theme_minimal()

```

```{r}
#| echo: false
#| warning: false
#| label: fig-swharvesthunters
#| fig-cap: "Number of active bear hunters plotted against total bear harvests from 2012-2023 in the southwest region"

#create line graph 
ggplot()+
  geom_line(data = sw_bear, aes(x = `Active Hunters`, y = Harvest), color = "blue") +
  theme_minimal()
```

```{r}
#| echo: false
#| warning: false
#| label: fig-seharvesthuntersbear
#| fig-cap: "Number of active bear hunters plotted against total bear harvests from 2012-2023 in the southeast region"

#create line graph 
ggplot()+
  geom_line(data = se_bear, aes(x = `Active Hunters`, y = Harvest), color = "blue") +
  theme_minimal()
```

```{r}
#| echo: false
#| warning: false
#| label: fig-northharvesthuntersbear
#| fig-cap: "Number of active bear hunters plotted against total bear harvests from 2012-2023 in the northern region"

#create line graph 
ggplot()+
  geom_line(data = north_bear, aes(x = `Active Hunters`, y = Harvest), color = "blue") +
  theme_minimal()
```

```{r}
#| echo: false
#| warning: false
#| label: fig-swharvesthuntersmoose
#| fig-cap: "Number of active moose hunters plotted against total moose harvests from 2009-2023 in the southwest region"

#create line graph 
ggplot()+
  geom_line(data = sw_moose, aes(x = `Active Hunters`, y = `Total Harvest`), color = "red") +
  theme_minimal()
```

```{r}
#| echo: false
#| warning: false
#| label: fig-seharvesthuntersmoose
#| fig-cap: "Number of active moose hunters plotted against total moose harvests from 2009-2023 in the southeast region"

#create line graph 
ggplot()+
  geom_line(data = se_moose, aes(x = `Active Hunters`, y = `Total Harvest`), color = "red") +
  theme_minimal()
```

```{r}
#| echo: false
#| warning: false
#| label: fig-northharvesthuntersmoose
#| fig-cap: "Number of active moose hunters plotted against total moose harvests from 2009-2023 in the northern region"

#create line graph 
ggplot()+
  geom_line(data = north_moose, aes(x = `Active Hunters`, y = `Total Harvest`), color = "red") +
  theme_minimal()
```

# Model

# Results

# Discussion

## What was done

## Something we learned about the world

## Another thing we learned about the world

## Weaknesses

## Looking forward

# References